---
title: "Statistical Inference Course Project"
author: "Anton Titov"
date: '5-April-2016 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, cache=FALSE,  warning=FALSE, results="hide")
options(scipen = 1, digits = 4)
```
## Intro
This report will cover the analysis of exponential distribution. We will compare parameters of sample distribution and theoretical one. 
## Loading and preprocessing the data
We are loading **data.table** for fast data handling and **ggplot2**. 
Also we change system time display to English by **Sys.setlocale**.
```{r Locale and Libraries}
Sys.setlocale("LC_TIME", "English")
library(data.table)
library(ggplot2)
set.seed(1234521)
```
## Playing with exponential distribution
We set parameters for exponential distributions. 
Then we generate vector of means for 50 samples.
After that we calculate sample mean and sample standart deviation.
```{r}
smean <- sexpo <- NULL
lambda <- 0.2
nexp <- 50
n <- 1000
for (i in 1:n) 
  {
  expounit <- rexp(nexp, lambda)
  sexpo <- c(sexpo, expounit)
  #smean <- c(smean, mean(rexp(nexp, lambda)))
  smean <- c(smean, mean(expounit))
  
}
simumean <- mean(smean)
simusd <- sd(smean)
```
After that we simulate variance.
```{r}
var <- apply(matrix(rexp(n * nexp, lambda), n), 1, var)
simuvar <- mean(var)
```
We create data.frame of sample and theoretical means and variances for easy plotting
```{r}
bothmeans <- data.frame(Mean=c("Sample","Theoretical"), vals = c(simumean,1/lambda))
bothvars <- data.frame(Var=c("Sample","Theoretical"),vals = c(simuvar,(1/lambda)^2))
```

## Question 1. 
###Show the sample mean and compare it to the theoretical mean of the distribution

```{r}
g = ggplot(data.frame(smean = smean), aes(smean))
g = g + geom_histogram(aes(y = ..density..), fill = "lightgrey", colour = "black", bins=50) 
g = g +  geom_vline(data = bothmeans, aes(xintercept = vals, color = Mean, linetype = Mean), size = 1)
g = g +  theme(panel.background = element_rect(fill = 'white'), legend.position = "bottom")
g = g + labs(x="Mean", y="Density", title="50 samples of Exponential distribution mean")
print(g)
```
```{r, echo= FALSE}
dev.off()
```
As we can see sample mean is pretty close to theoretical.  
Sample mean: `r simumean`  
Theoretical mean: `r 1/lambda`  
Sample SD of the mean: `r simusd` 

## Question 2. 
###Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.


```{r}
g = ggplot(data.frame(var = var), aes(var))
g = g + geom_histogram(aes(y = ..density..), fill = "lightblue", color = "black", bins=50) + 
  stat_function(fun = dnorm, size = 2, args = list(mean = (1/lambda)^2, sd = sd(var)))
g = g + geom_vline(data = bothvars, aes(xintercept = vals, color = Var, linetype = Var),size = 1)
print(g)
```
```{r, echo= FALSE}
dev.off()
```
As we can see sample variance is pretty close to theoretical.  
Sample varinace: `r var(smean)`  
Theoretical variance: `r (1/lambda)^2/nexp`  
Sample SD of the mean: `r sd(smean)`  
Theoretical SD of the mean `r (1/lambda)/sqrt(nexp)`  
## Question 3.
###Show that the distribution is approximately normal.

```{r}
g = ggplot(data.frame(smean = smean), aes(smean))
g = g + geom_histogram(aes(y = ..density..), fill = "lightgrey", colour = "black", bins=50) + 
      stat_function(fun = dnorm, size = 1, args = list(mean = simumean, sd = simusd), aes(color="Theoretical density") ) +
      geom_density( size=1, aes(color="Simulation density"), linetype = 2) +
      scale_color_manual(values=c("Simulation density" = "blue", "Theoretical density" = "red"), 
                        name="Experimental vs Theroretical\nComparison") +
      guides(colour = guide_legend(override.aes = list(linetype=c(2,1))), 
             fill = guide_legend(override.aes = list(colour = NULL)))
      
g = g + geom_vline(xintercept = simumean, size = 1)
g = g +  theme(panel.background = element_rect(fill = 'white'), legend.position = "bottom")
g = g + labs(x="Mean", y="Density", title="50 samples of Exponential distribution mean")
print(g)
```
```{r, echo= FALSE}
dev.off()
```
Distribution of the sample mean looks like normal. If we increase nubmer of experiments then it will look more normal. 